---
title: "Tidy text with R"
author: "Rahadian Arthapati"
date: "10/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Tidy text
Tidy data as described by Hadley Wickham, has specific structure :

- Each variable is a column
- Each observation is row
- Each type of observational unit is a table

We thus define the tidy text format as being a table with one token per row. A token is a meaningful unit of text, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.

#References
Julia Silge. *Text Mining with R: A Tidy Approach.* 

```{r}
#Let’s use the text of Jane Austen’s 6 completed, published novels from the janeaustenr package, and transform them into a tidy format.
library(janeaustenr)
library(tidytext)
library(tidyverse)
library(glue)
library(stringr)
library(ggplot2)
```

```{r}
(original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = T)))) %>%
  ungroup())
```
To work with this as a tidy dataset, we need to restructure it as one-token-per-row format. The unnest_tokens function is a way to convert a dataframe with a text column to be one-token-per-row
```{r}
(tidy_books <- original_books %>%
  unnest_tokens(word, text))
```
We can remove stop words (kept in the tidytext dataset stop_words) with an anti_join.We can also use count to find the most common words in all the books as a whole.
```{r}
data("stop_words")
(cleaned_books <- tidy_books %>%
  anti_join(stop_words) %>% count(word, sort=T))
```
We can create visualization of the most common words
```{r}
cleaned_books%>%
  filter(n>600)%>%
  mutate(word=reorder(word,n))%>%
  ggplot(aes(word,n))+geom_col()+xlab(NULL)+ylab("count")+coord_flip()
```
Now that we’ve used the janeaustenr package to explore tidying text, let’s introduce the gutenbergr package (Robinson 2016). The gutenbergr package provides access to the public domain works from the Project Gutenberg collection. The package includes tools both for downloading books (stripping out the unhelpful header/footer information), and a complete dataset of Project Gutenberg metadata that can be used to find works of interest.

Let’s look at some science fiction and fantasy novels by H.G. Wells, who lived in the late 19th and early 20th centuries. Let’s get The Time Machine, The War of the Worlds, The Invisible Man, and The Island of Doctor Moreau. We can access these works using gutenberg_download() and the Project Gutenberg ID numbers for each novel.
```{r}
library(gutenbergr)

hgwells <- gutenberg_download(c(35, 36, 5230, 159))

tidy_hgwells <- hgwells %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

tidy_hgwells%>%count(word,sort=T)
```
Now let’s get some well-known works of the Brontë sisters, whose lives overlapped with Jane Austen’s somewhat, but who wrote in a rather different style. Let’s get Jane Eyre, Wuthering Heights, The Tenant of Wildfell Hall, Villette, and Agnes Grey.
```{r}
bronte <- gutenberg_download(c(1260, 768, 969, 9182, 767))

tidy_bronte <- bronte %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

tidy_bronte %>%
  count(word, sort = T)
```
Now, let’s calculate the frequency for each word in the works of Jane Austen, the Brontë sisters, and H.G. Wells by binding the data frames together. We can use spread and gather from tidyr to reshape our data frame so that it is just what we need for plotting and comparing the three sets of novels.
```{r}
#We use str_extract() here because the UTF-8 encoded texts from Project Gutenberg have some examples of words with underscores around them to indicate emphasis (like italics). The tokenizer treated these as words, but we don’t want to count “any” separately from “any” as we saw in our initial data exploration before choosing to use str_extract().

(frequency <- bind_rows(mutate(tidy_bronte, author = "Brontë Sisters"),
                       mutate(tidy_hgwells, author = "H.G. Wells"),
                       mutate(tidy_books, author = "Jane Austen")) %>%
  mutate(word = str_extract(word, "[a-z']+"), author=gsub("\303\253","e",author)) %>%
  count(author, word) %>%
  group_by(author) %>%
  mutate(proportion = n / sum(n)) %>%
  select(-n) %>%
  spread(author, proportion) %>%
  gather(author, proportion, `Bronte Sisters`:`H.G. Wells`))

```
Let's plot
```{r}
library(scales)
# expect a warning about rows with missing values being removed
ggplot(frequency, aes(x = proportion, y = `Jane Austen`,
                      color = abs(`Jane Austen` - proportion))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3)+
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001),
                       low = "darkslategray4", high = "gray75") +
  facet_wrap(~author, ncol = 2) +
  theme(legend.position="none") +
  labs(y = "Jane Austen", x = NULL)
```
Overall, notice in plot above that the words in the Austen-Brontë panel are closer to the zero-slope line than in the Austen-Wells panel. Also notice that the words extend to lower frequencies in the Austen-Brontë panel; there is empty space in the Austen-Wells panel at low frequency. These characteristics indicate that Austen and the Brontë sisters use more similar words than Austen and H.G. Wells. Also, we see that not all the words are found in all three sets of texts, and there are fewer data points in the panel for Austen and H.G. Wells.

Let’s quantify how similar and different these sets of word frequencies are using a correlation test. How correlated are the word frequencies between Austen and the Brontë sisters, and between Austen and Wells?
```{r}
#If the p-value is < 5%, then the correlation between x and y is significant.
cor.test(data = frequency[frequency$author == "Bronte Sisters",],
         ~ proportion + `Jane Austen`)

cor.test(data = frequency[frequency$author == "H.G. Wells",],
         ~ proportion + `Jane Austen`)
```























